A Type System for Safe and Scalable Parallelism using Logical Regions \\ (Extended Version)

### 2012-07-20 doc/popl2013full/langdef.tex
few more edits, one new proof to coherence section

The proof of this (and other lemmas in this section) is included in \cite{LegionTypes12}.
We now show that the concatenation of the memory traces from these serially
executed subexpressions is always a valid interleaving:

@@

\begin{proof}
First, we must show that $any\_interleave(E', \oton{E})$.  This is done by
induction on the length of $E_1$, followed by $E_2$, and so on.  Next, we must
show that $E'$ is coherent for both $L_{excl}$ and $L_{atomic}$.  This is done by
matching the structure of $E'$ with $\oton{E}$ and the use of the following
lemma: $apply(S, E_1 \dplus E_2) = apply(apply(S, E_1), E_2)$.  With the
question of coherence addressed, the sequential equivalence of $E'$ is trivial
for both the exclusive and atomic cases.
\end{proof}

@@

\begin{proof}
The proof is by contradiction.  Assume $E_1 \nonint E_2$.  Then there must be some $\epsilon_1$ in $E_1$ and some $\epsilon_2$ in $E_2$ such that $\epsilon_1 \nonint \epsilon_2$.  This requires that
$\epsilon_1$ and $\epsilon_2$ be operations to the same location $l$ and not both be reads or both
reductions using the same function name.  Assume $\epsilon_1 = read(l, c_1, v_1)$ and $\epsilon_2 = write(l, c_2, v_2)$.  The definition of $\privconsist{E_1}{\Phi_1}$ tells us that there must be some $r_1$ satisfying $l \in M(r_1) \wedge reads(r_1) \in \Phi_1$.  Similarly, there must be some $r_2$ satisfying $l \in M(r_2) \wedge writes(r_2) \in \Phi_2$.  Letting $\phi_1 = reads(r_1)$ and $\phi_2 = writes(r_2)$, we have $M(r_1) \cap M(r_2) \not= \emptyset$ (it contains at least $l$), so $\phi_1 {\nonint[M]} \phi_2$ and $\Phi_1 {\nonint[M]} \Phi_2$.  There are seven other cases to consider (using $id_1 \not= id_2$ for the reduce-reduce case), but all yield the same result.
\end{proof}

>>>>

\begin{proof}
The proof is by induction on the structure of the derivation of $e$.  The clobber set is only manipulated
in one case and only examined in one other case.  In all other cases, the clobber set is unchanged and
used only in the derivation of subexpressions.  The case where the clobber set is manipulated is [E-Call], but
the use of a union operation still allows the inductive hypothesis to be applied to the subexpression $e_{n+1}$.
The examination of the clobber set occurs in [E-Read] where it is used to choose the result $v$ of the read
from the heap.  For each use of the [E-Read] rule, there are three possible cases:
\begin{itemize}
\item $l \in C$, for which $v = S'(l)$ in the derivation using $C$.  Since $C \subseteq C'$, we have $l \in C'$, and $S'(l)$ is the correct result for the $C'$ derivation as well.
\item $l \not\in C'$, and therefore $l \not\in C$.  Both derivations may use any value $v' : H(l)$.
\item $l \in C, l \not\in C'$, for which $v = S'(l)$ must be chosen for the derivation using $C$.  From Theorem~\ref{thm:effects} and Lemma~\ref{lemma:heapconst:apply}, we have $\traceconsist{E}$ and $\storeconsist{S' = apply(S, E)}{H}$, so $\valueconsist[]{S'(l)}{H(l)}$, and $v$ is a legal choice for $v'$ in the derivation using $C'$.
\end{itemize}
\end{proof}

For the following proofs, we will need one additional lemma, which follows directly
from the definition of $apply$:

\begin{lem}
\label{lemma:applychain}
\rm
Let $S$ be an initial store, and $E_1$ and $E_2$ be memory traces.  Then:
$$apply(S, E_1 \dplus E_2) = apply(apply(S, E_1), E_2)$$
\end{lem}

We now show that the concatenation of the memory traces from these serially
executed subexpressions is always a valid interleaving:

@@

\begin{proof}
First, we must show that $any\_interleave(E', \oton{E})$.  This is done by
induction on the length of $E_1$, followed by $E_2$, and so on.  Next, we must
show that $E'$ is coherent for both $L_{excl}$ and $L_{atomic}$.  This is done by
matching the structure of $E'$ with $\oton{E}$ and the use of Lemma~\ref{lemma:applychain}.
With the
question of coherence addressed, the sequential equivalence of $E'$ is trivial
for both the exclusive and atomic cases.
\end{proof}

@@

\begin{proof}
The proof is by contradiction.  Assume $E_1 \cancel\nonint E_2$.  Then there must be some $\epsilon_1$ in $E_1$ and some $\epsilon_2$ in $E_2$ such that $\epsilon_1 \cancel\nonint \epsilon_2$.  This requires that
$\epsilon_1$ and $\epsilon_2$ be operations to the same location $l$ and not both be reads or both
reductions using the same function name.  Assume $\epsilon_1 = read(l, c_1, v_1)$ and $\epsilon_2 = write(l, c_2, v_2)$.  The definition of $\privconsist{E_1}{\Phi_1}$ tells us that there must be some $r_1$ satisfying $l \in M(r_1) \wedge reads(r_1) \in \Phi_1$.  Similarly, there must be some $r_2$ satisfying $l \in M(r_2) \wedge writes(r_2) \in \Phi_2$.  Letting $\phi_1 = reads(r_1)$ and $\phi_2 = writes(r_2)$, we have $M(r_1) \cap M(r_2) \not= \emptyset$ (the intersection contains at least $l$), so $\phi_1 \cancel{\nonint[M]} \phi_2$ and $\Phi_1 \cancel{\nonint[M]} \Phi_2$.  There are seven other cases to consider (using $id_1 \not= id_2$ for the reduce-reduce case), but all yield the same result.
\end{proof}
